### 2023-06-13 Complete-V1/Complete-V1.tex

% !TeX program = pdfLaTeX
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}

%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

%% load any required packages here

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\begin{document}

\title{Machine Learning Deriven Set of microRNAs as a Novel Biomarker
for Myocardial Infarction Diagnosis }

\author{  Mehrdad Samadishadlou \and  Reza Rahbarghazi \and  Farhad
Bani \and  }

    \authorrunning{ M. Samadishadlou et al. }

\institute{
        Mehrdad Samadishadlou \at
     Medical Nanotechnology gruop, Tabriz University of Medical
Sciences, Tabriz, Iran \\
     \email{\href{mailto:samadishadlou@tbzmed.ac.ir}{\nolinkurl{samadishadlou@tbzmed.ac.ir}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Reza Rahbarghazi \at
     Medical Nanotechnology gruop, Tabriz University of Medical
Sciences, Tabriz, Iran \\
     \email{\href{mailto:rahbarghazir@tbzmed.ac.ir}{\nolinkurl{rahbarghazir@tbzmed.ac.ir}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Farhad Bani \at
     Medical Nanotechnology gruop, Tabriz University of Medical
Sciences, Tabriz, Iran \\
     \email{\href{mailto:bainf@tbzmed.ac.ir}{\nolinkurl{bainf@tbzmed.ac.ir}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
    }

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
MicroRNAs (miRNAs) play a crucial role in regulating adaptive and
maladaptive responses in cardiovascular diseases, making them attractive
targets for potential therapeutics. However, their potential as novel
biomarkers for diagnosing cardiovascular diseases requires systematic
evaluation. In this study, we aimed to identify a key set of miRNA
biomarkers using integrated bioinformatics and machine learning
analysis. We combined and analyzed three gene expression datasets from
the Gene Expression Omnibus (GEO) database, which contains peripheral
blood mononuclear cell samples from individuals with myocardial
infarction (MI), stable coronary artery disease (CAD), and healthy
individuals. Additionally, we selected a set of miRNAs based on their
area under the curve of the receiver operating characteristic curve
(AUC-ROC) to separate the CAD/MI samples. We designed a two-layer
architecture for sample classification, with the first layer isolating
healthy samples from not-healthy ones, and the second layer classifying
stable CAD and MI samples. We trained different machine learning models
using both biomarker sets and evaluated their performance on a test set.
We identified miR-21, miR-186, and miR-32 as the only miRNAs among the
differentially expressed genes, and a set including miR-186, miR-21,
miR-197, miR-29A, and miR-296 as the optimum set of miRNAs selected by
their AUC-ROC. Both biomarker sets could distinguish healthy from
not-healthy samples with complete accuracy. The best performance for the
classification of CAD and MI was achieved with an SVM model trained
using the biomarker set selected by AUC-ROC, with an AUC-ROC of 0.95,
and an accuracy of 0.88. Our study demonstrated that miRNA signatures
derived from peripheral blood could serve as valuable novel biomarkers
for cardiovascular diseases.
\\
\keywords{
        microRNA \and
        Machine Learning \and
        Myocardial Infarction \and
    }

\end{abstract}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

At present, cardiovascular diseases (CVDs) are the leading cause of
human mortality with 32\% of all global deaths. It is estimated that
about 85\% of CVDs mortality were diagnosed with myocardial infarction
(MI) ({``Cardiovascular Diseases ({CVDs})''} n.d.). MI is an acute
coronary syndrome with sudden blockage and stenosis of the coronary
artery, and subsequent myocardial ischemia, leading to extensive
cardiomyocyte damage and necrosis (Yap et al. 2023).

Over the last 50 years, numerous attempts have been collected to use
biomarkers to facilitate diagnosis, assess risk, follow-up therapy, and
determine therapeutic efficacy in CVDs candidates. Based on the released
guidelines, cardiac troponins (cTns) are used as a highly-sensitive and
accurate approach for the detection of MI. Despite the inherent
advantages, the high-rate sensitivity of cTn-based assays has also led
to more false positive results (Thygesen et al. 2018), which do
necessitate the advent and development of new modalities with
pathological values. To improve diagnostic value upon existing MI
biomarkers, the combination of complementary biological markers, such as
microRNAs (miRNAs) and other genetic factors, is proposed. Previous
researches support the notion that miRNAs exhibit the great potential to
be used as alternative biomarkers in CVDs detection and follow-up
(Schulte et al. 2020). It is suggested that miRNAs possess 18-22
nucleotides and can play a crucial role in the regulation of gene
expression. Evidence point to the fact that miRNAs are involved in the
pathogenesis of cardiac tissue injury and could be considered as
theranostics targets in terms of CVDs. These elements can easily
circulate in biofluids with both diagnostic as well as prognostic values
(Schulte, Karakas, and Zeller 2017). Several biological activities such
as angiogenesis, cardiomyocyte growth and contractility, lipid
metabolism, plaque formation, and cardiac rhythm are regulated by miRNAs
(Kalayinia et al. 2021). It is postulated that the function and
diagnostic properties of miRNAs are beyond the myocardium in CVD
patients. To be specific, the expression of miRNAs can vary in different
biofluids and cell components such as serum and peripheral blood
mononuclear cells (PBMCs) (Soler-Botija, Gálvez-Montón, and Bayés-Genís
2019).

PBMCs are a fraction of white blood cells (WBCs), including monocytes,
lymphocytes, macrophages, and other cells belonging to the immune system
(Gao et al. 2020). Emerging data have indicated that PBMCs can be used
as a valid source of biomarkers for monitoring various pathological
conditions. Of note, the alteration of mRNAs and miRNAs under
pathological conditions gives us valuable information about different
kinds of disorders. PBMCs could recapitulate the conditions of the
target tissues, thus, providing a highly sensitive and specific source
of biomarkers (Mosallaei et al. 2022). Commensurate with these
conditions, these cells are repositories of dysregulated genes and
miRNAs expression profiles in CVDs (Gao et al. 2020; Mosallaei et al.
2022).

In recent years, the advent and use of machine learning (ML) is an
exciting prospect for advancing scientific discoveries. Although the
concept of ML and its initial algorithms were conceived many years ago,
recent improvements in computing power and access to vast amounts of
data have shown that ML techniques outperform classical statistical
methods in various fields. Furthermore, the progress made in omics
technologies has enabled the analysis of massive and intricate
biological data sets, consisting of hundreds to thousands of samples,
which makes it possible for ML to extract valuable biological
information from such data (Torun et al. 2023). Therefore, ML offers
novel techniques to integrate and analyze the various omics data
enabling the discovery of de novo biomarkers. These biomarkers help us
in accurate disease prediction, patient stratification, and finding new
therapeutics (Reel et al. 2021).

In this study, we aimed to identify potential miRNA biomarkers for MI
patients by combining and analyzing three different microarray datasets
from PBMCs. It is suggested that the integration of omics data with
bioinformatics and ML techniques could be a promising tool in the
discovery of new and more accurate biomarkers for monitoring MI.
Besides, this approach can deepen our vision into the underlying
mechanisms of MI and aid in the development of valid diagnostic
biomarkers, and patient stratification.

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{microarray-data-collection}{%
\subsection{Microarray data
collection}\label{microarray-data-collection}}

Microarray datasets were obtained from the Gene Expression Omnibus (GEO)
database (\url{https://www.ncbi.nlm.nih.gov/geo/}). To obtain sufficient
classification power between MI, healthy and stable coronary artery
disease (CAD) samples, a relatively large sample size was required.
Therefore, GSE59867 for MI and CAD samples, and GSE56609 and GSE54475
for healthy samples were selected. All samples were produced using
Affymetrix Human Gene 1.0 ST Array (GPL6244) platform. Only healthy,
stable CAD and early-stage MI samples were selected from these datasets
for further analyses. The basic information for the three datasets
evaluated in the current study is provided in Table @ref(tab:datasets).

\begin{longtable}[]{@{}cccccc@{}}
\caption{Basic information of the GEO microarray
datasets.}\tabularnewline
\toprule()
Dataset & Platform & Healthy & CAD & MI & Refrence \\
\midrule()
\endfirsthead
\toprule()
Dataset & Platform & Healthy & CAD & MI & Refrence \\
\midrule()
\endhead
GSE59867 & GPL6244 & - & 46 & 111 & (Maciejak et al. 2015) \\
GSE56609 & GPL6244 & 46 & - & - & (Matone et al. 2015) \\
GSE54475 & GPL6244 & 5 & - & - & (Canali et al. 2014) \\
\bottomrule()
\end{longtable}

\hypertarget{pre-processing}{%
\subsection{Pre-processing}\label{pre-processing}}

Raw data (CEL files) of all datasets were downloaded from the GEO and
pre-processed using the fRMA package (M. N. McCall, Bolstad, and
Irizarry 2010). fRMA allowed to pre-process of individual microarray
samples and combining them consistently for analysis. For each dataset,
background correction was performed using the RMA algorithm and then it
was quantile normalized based on the reference distribution. During
summarization, batch effects were removed and variances of the gene
expressions were estimated by taking into account these probe-specific
effects. For those multiple probe sets matched to the identical gene,
the mean log fold change was retained. Therefore, fRMA can be seen as a
batch effect removal technique for different datasets that are produced
by identical microarray platforms. Thus, to ensure batch effect removal,
the principal component analysis and the relative log expression of
train samples were plotted before and after fRMA (Lazar et al. 2013).

\hypertarget{differential-expression-analysis}{%
\subsection{Differential expression
analysis}\label{differential-expression-analysis}}

The barcode algorithm proposed by McCall et al. (Matthew N. McCall et
al. 2011) transformed the actual expression values into binary barcode
values. Huge sets of samples were collected and normalized using fRMA
for several platforms as well as for Affymetrix Human Gene 1.0 ST Array
(GPL6244) platform. The distribution of the expressed and unexpressed
observed intensities for each gene is estimated using these normalized
sets. Genes were considered expressed (and their value coded to 1) or
unexpressed (and their value coded to 0) according to the following
equation:

\[
\hat{x}_{ij} = \left\{
  \begin{array}{ll}
    1 & \mbox{if } x_{ij} >= \mu^{ne} + C \times \sigma^{ne} \\
    0 & \mbox{otherwise}
  \end{array}
\right.
\]

where \(x_{ij}\) is the normalized intensity of gene \(i\) in sample
\(j\), \(C\) is a user-defined parameter, \(\sigma^{ne}\) is the
standard deviation of the non-expressed distribution, and \(\mu^{ne}\)
is the mean of the non-expressed distribution. The barcode
representation of a sample is a vector of ones and zeros denoting which
genes are estimated to be expressed (ones) and unexpressed (zeros). The
barcode algorithm was implemented by the barcode function in the R fRMA
package, and the default value of \(C\) was used.

To determine if the expressed ratios differed in the MI group versus the
healthy control group, Fisher's exact test for individual genes was
carried out upon the barcode values. Genes with a false discovery rate
(FDR) of \(< 0.05\), which was calculated through the Benjamini-Hochberg
(BH) procedure to adjust for multiple testing issues, were considered as
differentially expressed genes. The same procedures were conducted on
CAD versus healthy controls as well as MI versus CAD group to find the
DEGs between them.

\hypertarget{functional-and-pathway-enrichment-analyses}{%
\subsection{Functional and pathway enrichment
analyses}\label{functional-and-pathway-enrichment-analyses}}

Using the R clusterProfiler package, the Kyoto Encyclopedia of Genes and
Genomes (KEGG) pathway enrichment analysis and Gene Ontology (GO)
functional annotation were carried out on the differentially expressed
genes. The GO analysis included biological process (BP), cellular
component (CC) and molecular function (MF) categories. An adjusted
p-value of less than 0.05 was considered to indicate a statistically
significant difference. Enrichments were conducted on the MI-healthy and
CAD-healthy DEGs. In these analyses, all default parameters were used.

\hypertarget{ml-procedure}{%
\subsection{ML procedure}\label{ml-procedure}}

The ML analysis was performed using Python software, ver. 3.9, Numpy
(Harris et al. 2020), pandas (McKinney 2010), and Scikit-Learn packages
(Pedregosa et al. 2011). Whenever hyper-tuning was needed, the
scikit-opt package (Head et al. 2021) was used. In all ML analyses, the
datasets were divided into train and test sets by a 0.7:0.3 ratio and
all reported results are the average of 10-fold cross-validation.

Two different approaches were used for selecting miRNAs for model
training. The first approach was using the miRNAs that are
differentially expressed. In the second approach, miRNAs with individual
area under the curve of the receiver operating characteristic (AUC-ROC)
over 0.8 for separating MI from CAD were selected. Having the result of
these two different approaches can provide an informative comparison
between the predictive capabilities of sets of miRNAs selected with
different logics.

\hypertarget{mirnas-in-degs}{%
\subsubsection{miRNAs in DEGs}\label{mirnas-in-degs}}

In this approach, a two-layer architecture was deployed to the data to
maximize the prediction values. The first layer predicted whether a
sample is healthy or not, and the second layer separated MI from CAD in
the samples which were predicted as not healthy in the first layer. To
this end, a distinct ML model was trained for each layer. Since there is
a limited number of miRNAs in DEGs, both layers were trained with all of
them. For further comparison with the models' performance, the ROC curve
of each miRNA for classifying healthy and not-healthy, as well as CAD
and MI, were generated using a Logistic Regression model.

\hypertarget{first-layer-for-isolation-of-healthy-and-not-healthy-samples}{%
\paragraph{First layer for isolation of healthy and not-healthy
samples:}\label{first-layer-for-isolation-of-healthy-and-not-healthy-samples}}

An SVM model using RBF kernels was trained and hyper-tuned using all
miRNAs in DEGs. To handle the severe imbalance in the number of samples
(51 for the healthy and 157 for the not-healthy groups), the sample
weight for the healthy and the not-healthy samples were set to 1 and
0.5, respectively.

\hypertarget{second-layer-for-seperating-mi-and-cad-samples}{%
\paragraph{Second layer for seperating MI and CAD
samples:}\label{second-layer-for-seperating-mi-and-cad-samples}}

For the sake of reaching the highest classification performance using
the set of miRNAs, different models were investigated. To do so, SVM
(with linear, polynomial, and RBF kernels), Logistic Regression (LR),
Random Forests (RF), k-Nearest Neighbor (kNN), Gradient Boosting (GB),
XGBoost (XGB) and Decision Tree (DT) models were trained. All models
were trained with their pre-set parameters with 10-fold
cross-validation. The criteria for choosing the best model were the
highest accuracy and AUC-ROC on the test set. The best model was
hyper-tuned with the scikit-opt package (Head et al. 2021) to get the
best predictive performance.

\hypertarget{mirnas-with-the-highest-auc-roc}{%
\subsubsection{miRNAs with the highest
AUC-ROC}\label{mirnas-with-the-highest-auc-roc}}

Like the previous approach, a two-layers strategy was conducted. The
first layer classified samples into healthy and not-healthy, and the
separated MI and CAD samples. However, to keep the number of miRNAs as
low as possible miRNAs were selected from the second layer, and then
their performance wase evaluated in the first layer. AUC-ROC of all
miRNAs for classifying MI and CAD samples were calculated and the miRNAs
with the AUC-ROC over 0.8 were selected. The ROC curves for each
selected miRNA for separating healthy samples from not-healthy and MI
from CAD samples were also plotted for further comparison.

\hypertarget{first-layer-for-the-detection-of-healthy-and-not-healthy-samples}{%
\paragraph{First layer for the detection of healthy and not-healthy
samples:}\label{first-layer-for-the-detection-of-healthy-and-not-healthy-samples}}

An SVM model with an RBF kernel was trained using the selected set of
miRNAs. Additionally, the model was hyper-tuned to find the
hyper-parameters for the highest AUC-ROC and accuracy. The ROC curve and
confusion matrix for the best model were reported.

\hypertarget{second-layer-for-separating-mi-and-cad}{%
\paragraph{Second layer for separating MI and
CAD:}\label{second-layer-for-separating-mi-and-cad}}

The selected miRNAs set was used to train different algorithms to find
the best model. Similar to the previous approach, SVM (with linear,
polynomial, and RBF kernels), LR, RF, kNN, GB, XGB, and DT were trained.
All models were trained with their pre-set parameters using 10-fold
cross-validation. The models with the highest AUC-ROC and accuracy on
the test set were selected and hyper-tuned using the scikit-opt package
(Head et al. 2021). The ROC curve and confusion matrix for the best
model were reported.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{pre-processing-1}{%
\subsection{Pre-processing}\label{pre-processing-1}}

The PCA plots of the training samples are shown in Figure @ref(fig:PCA)A
and B. As shown, healthy samples were separated from CAD or MI samples
in primary data and also after conducting fRMA. In the RLE plot, there
was a distinct difference between dataset means for all samples before
conducting fRMA (Figure @ref(fig:PCA)C). All datasets were rearranged
around 0 in the RLE plot after conducting fRMA (Figure @ref(fig:PCA)D).
Moreover, there was a clear change in inter-quantile distances, but the
values still were over 0.1.

\begin{figure}
\includegraphics[width=1\linewidth]{../../DEG and DEM/Pics/PCA and RLE of trains} \caption{PCA and RLE plot for all samples before and after fRMA.}\label{fig:PCA}
\end{figure}

\hypertarget{differential-expression-analysis-1}{%
\subsection{Differential expression
analysis}\label{differential-expression-analysis-1}}

\begin{table}

\caption{\label{tab:DEGstab}Total, up-, and down-regulated DEGs and differentially expressed miRNAs.}
\centering
\begin{tabu} to \linewidth {>{\raggedright\arraybackslash}p{3cm}>{\centering}X>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{5cm}}
\toprule
  & Total DEGs & Up-regulated DEGs & Down-regulated DEGs & miRNAs\\
\midrule
MI vs. Healthy & 860 & 323 & 537 & hsa-miR-186, miR-21, miR-32\\
CAD vs. Healthy & 670 & 262 & 408 & hsa-miR-186, miR-21, miR-32\\
MI vs. CAD & 260 & 144 & 116 & hsa-miR-186\\
\bottomrule
\end{tabu}
\end{table}

According to the cutoff criterion of \(FDR < 0.05\), there were 860 DEGs
between the MI and the healthy samples. Among them, 323 were
up-regulated, and 537 were down-regulated in MI compared to the healthy
controls. In CAD and healthy groups comparison, we found 670 DEGs, of
which 262 and and 408 DEGs were up- and down-regulated, respectively in
CAD samples. In the MI and CAD groups, the number of DEGs was 260, and
the number of up- and down-regulated genes in MI samples were 144 and
116, respectively in comparison with CAD samples. These data are
summarized in Table @ref(tab:DEGstab).

\begin{figure}

{\centering \includegraphics[width=0.45\linewidth]{../../DEG and DEM/Pics/venny}

}

\caption{Venn diagram for DEGs in CAD/Healthy, MI/Healthy, and MI/CAD comparison.}\label{fig:venn}
\end{figure}

The Venn diagram in Figure @ref(fig:venn) shows that CAD and MI samples
share a majority of their DEGs. From 860 DEGs of MI/healthy and 670 DEGs
of CAD/healthy, 531 genes were common which is 62\% of MI/healthy DEGs
and 79\% of CAD/healthy DEGs.

\hypertarget{gene-ontology-go-and-kyoto-encyclopedia-of-genes-and-genomes-kegg-enrichment-analyses-of-the-degs.}{%
\subsection{Gene ontology (GO) and Kyoto Encyclopedia of Genes and
Genomes (KEGG) enrichment analyses of the
DEGs.}\label{gene-ontology-go-and-kyoto-encyclopedia-of-genes-and-genomes-kegg-enrichment-analyses-of-the-degs.}}

To explore the biological classification of the DEGs, we performed GO
and KEGG pathway enrichment analyses on MI-healthy and CAD-healthy DEGs.
In MI versus healthy samples, GO enrichment analysis in the BP category,
suggested that the DEGs were enriched in ``immune response-regulating
signaling pathway'', ``lymphocyte differentiation'', ``immune
response-regulating cell surface receptor signaling pathway'', and
``leukocyte activation involved in immune response'' (Figure
@ref(fig:MIHEnrich)A). In the CC category, the DEGs were enriched in
``secretory granule membrane'', ``azurophil granule'', ``ficolin-1-rich
granule'', ``tertiary granule'', and ``ficolin-1-rich granule membrane''
(Figure @ref(fig:MIHEnrich)B). In the MF category, the DEGs were
involved in ``cadherin binding'' and ``MHC class I protein binding''
(Figure @ref(fig:MIHEnrich)C). KEGG pathway analysis indicated that the
DEGs were related to the following pathways: ``Chemokine signaling
pathway'', ``Lipid and atherosclerosis'', and ``Hematopoietic cell
lineage'' (Figure @ref(fig:MIHEnrich)D).

\begin{figure}
\includegraphics[width=0.95\linewidth]{../../DEG and DEM/Pics/Enrichment MI-Healthy} \caption{Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the MI and healthy DEGs. (A) Biological process terms. (B) Cellular component terms. (C) Molecular function terms. (D) KEGG analysis.}\label{fig:MIHEnrich}
\end{figure}

The enrichment results for DEGs of CAD versus healthy samples were as
follows. In the BP category, GO enrichment suggested that the DEGs were
enriched in ``positive regulation of defense response'', ``positive
regulation of innate immune response'', ``mononuclear cell
differentiation'', and ``positive regulation of response to external
stimulus'' (Figure @ref(fig:CADHEnrich)A). In the CC category, the DEGs
were enriched in ``azurophil granule'', ``ficolin-1-rich granule'', and
``ficolin-1-rich granule membrane'' (Figure @ref(fig:CADHEnrich)B). In
the MF category, the DEGs were involved in ``lipoprotein particle
receptor binding'' and ``NF-\(\kappa\)B binding'' (Figure
@ref(fig:CADHEnrich)C). KEGG pathway analysis indicated that the DEGs
were related to the following pathways: ``Chemokine signaling pathway'',
``Lipid and atherosclerosis'', and ``Hematopoietic cell lineage''
(Figure @ref(fig:CADHEnrich)D).

\begin{figure}
\includegraphics[width=0.95\linewidth]{../../DEG and DEM/Pics/Enrichment CAD-Healthy} \caption{Gene Ontology (GO) and Kyoto Encyclopedia of Genes and Genomes (KEGG) pathways enriched with the CAD and healthy DEGs. (A) Biological process terms. (B) Cellular component terms. (C) Molecular function terms. (D) KEGG analysis.}\label{fig:CADHEnrich}
\end{figure}

\hypertarget{machine-learning}{%
\subsection{Machine Learning}\label{machine-learning}}

\hypertarget{mirnas-in-degs-1}{%
\subsubsection{miRNAs in DEGs}\label{mirnas-in-degs-1}}

Among all DEGs, just miR-186, miR-32, and miR-21 were detected as
differentially expressed miRNAs. The expression profile of these three
miRNAs is presented in Figure @ref(fig:Expall). Additionally, The ROC
curves of each miRNA for each layer are presented in Figure
@ref(fig:miRROC). Using the logistic regression model the AUC-ROC of
miR-21, miR-32, and miR-186 for separating healthy and not-healthy
samples was 0.98, 0.99, and 0.90 respectively (Figure
@ref(fig:miRROC)A). Besides, the accuracy of each miRNA for classifying
the samples into healthy and not-healthy groups on the test set was
0.92, 0.98, and 0.89 for miR-21, miR-32, and miR-186, respectively.
Moreover, the ROC curve of each miRNA for classifying MI and CAD samples
was presented in Figure @ref(fig:miRROC)B. For miR-21, miR-32, and
miR-186, the AUC-ROC and accuracy on the test set were 0.85; 0.70; and
0.86, and 0.78; 0.67; and 0.74, respectively.

\begin{table}

\caption{\label{tab:mirExptable}Target miRNAs log fold-change and adjusted p-values for CAD samples relative to healthy, MI samples relative to healthy, and MI samples relative to CAD.}
\centering
\begin{tabu} to \linewidth {>{\raggedright}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{CAD/Healthy} & \multicolumn{2}{c}{MI/Healthy} & \multicolumn{2}{c}{MI/CAD} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7}
  & logFC & adj. p-value & logFC & adj. p-value & logFC & adj. p-value\\
\midrule
miR-186 & 1.4 & 3.60e-25 & 0.9 & 6.76e-20 & -0.5 & 1.05e-09\\
miR-21 & 1.4 & 1.31e-17 & 2.3 & 2.07e-47 & 0.8 & 2.96e-11\\
miR-32 & 2.5 & 8.39e-43 & 2.2 & 3.10e-59 & -0.3 & 7.60e-04\\
miR-197 & 0.5 & 2.95e-20 & 0.7 & 1.59e-47 & 0.2 & 8.58e-09\\
miR-29A & 0.7 & 7.76e-29 & 0.1 & 1.70e-01 & -0.5 & 2.14e-10\\
miR-296 & -0.1 & 5.00e-02 & 0.1 & 2.00e-02 & 0.2 & 6.15e-06\\
\bottomrule
\end{tabu}
\end{table}

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../../ML/Pics/Expressionforall}

}

\caption{Expression profile of all miRNAs in two approach in different sample classes.}\label{fig:Expall}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{../../ML/Pics/miRs ROCs}

}

\caption{ROC curve for single miRNAs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples.}\label{fig:miRROC}
\end{figure}

\hypertarget{first-layer-for-healthy-not-healthy-isolation}{%
\paragraph{First layer for healthy not-healthy
isolation:}\label{first-layer-for-healthy-not-healthy-isolation}}

Although single miRNAs had acceptable performance, their predictive
value could be improved even further by using them as a set. The ROC
curve for the SVM model with an RBF kernel trained with all three miRNAs
is presented in Figure @ref(fig:DEMROC)A. The model had a better
performance in classification than single miRNAs.The AUC-ROC for the
model is 1, and its accuracy on the test set was also 1. In Figure
@ref(fig:DEMCF)A, the confusion matrix for the model is presented.

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{../../ML/Pics/DEMs ROCs h not h cad mi}

}

\caption{ROC curve for miRNAs in DEGs on test set classification for (A) healthy and not-healthy samples and (B) CAD and MI samples.}\label{fig:DEMROC}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../../ML/Pics/DEMs Conf. matrix}

}

\caption{Confusion matrix on the test set for (A) An SVM model with RBF kernel for healthy and not-healthy and (B) An SVM model with linear kernel for CAD and MI samples classification.}\label{fig:DEMCF}
\end{figure}

\hypertarget{second-layer-for-separating-mi-samples-from-cad}{%
\paragraph{Second layer for separating MI samples from
CAD:}\label{second-layer-for-separating-mi-samples-from-cad}}

Different models were trained using expression values for three
differentially miRNAs. The models' 10-fold cross-validated AUC-ROC and
accuracy on the test set are reported in Figure @ref(fig:DEMmodels). The
best model from both AUC-ROC and accuracy point-of-view was the SVM
model with linear kernel. The AUC-ROC and accuracy for this model with
its pre-set values were 0.93 and 0.82 respectively. The model was
hyper-tuned for C and gamma hyper-parameters, and therefore the model
showed better performance. The ROC curve of the hyper-tuned model is
presented in Figure @ref(fig:DEMROC)B. For this model the AUC-ROC
reached 0.95 and the accuracy improved to 0.85 (Table @ref(tab:DEGsML)).
Moreover, the sensitivity and specificity for the model on the test set
were 0.91 and 0.71 respectively. The confusion matrix for the
hyper-tuned model is illustrated in Figure @ref(fig:DEMCF)B.

\begin{figure}

{\centering \includegraphics[width=0.65\linewidth]{../../ML/Pics/Models DEGs mirs}

}

\caption{Area under the receiver operating characteristic curve and accuracy of different models trained with three miRNAs in DEGs on the test set.}\label{fig:DEMmodels}
\end{figure}

\begin{table}

\caption{\label{tab:DEGsML}AUC-ROC and accuracy for SVM with the linear kernel as the best model trained with differentially expressed miRNAs on the train and test set before and after hyper-tuning}
\centering
\begin{tabu} to \linewidth {>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Pre-set parameters} & \multicolumn{2}{c}{Hyper-tuned} \\
\cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6}
Model & Metrics & train & test & train & test\\
\midrule
SVM-linear & AUC-ROC & 0.91 & 0.93 & 0.92 & 0.95\\
 & Accuracy & 0.83 & 0.82 & 0.84 & 0.85\\
\bottomrule
\end{tabu}
\end{table}

\hypertarget{auc-roc-approach}{%
\subsubsection{AUC-ROC approach}\label{auc-roc-approach}}

After calculating the AUC-ROC for each miRNA for the classification of
MI and CAD samples, the miRNAs with AUC-ROC over 0.8 were selected. The
selected miRNAs were miR-29a, miR-197, miR-186, miR-21, and miR-296. The
expression values of these miRNAs in healthy, CAD, and MI samples are
presented in Figure @ref(fig:Expall). The ROC curve of the selected
miRNAs for both layers are illustrated in Figure @ref(fig:miRROC).

\hypertarget{first-layer-for-isolation-of-healthy-and-not-healthy-samples-1}{%
\paragraph{First layer for isolation of healthy and not-healthy
samples}\label{first-layer-for-isolation-of-healthy-and-not-healthy-samples-1}}

Using the selected set, an SVM model with an RBF kernel was trained to
separate healthy from not-healthy samples. The ROC curve for the model
is presented in Figure @ref(fig:AUCROC)A and the confusion matrix is
illustrated in Figure @ref(fig:AUCCF)A. Both AUC-ROCand accuracy for the
model on the test set were equal to 1.

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{../../ML/Pics/AUC best model ROCs}

}

\caption{ROC curve for the set of miRNAs selected by AUC-ROC on test set classification. (A) SVM with RBF kernel for healthy and not-healthy samples classification. (B) Logistic regression model for CAD and MI samples classification. (C) SVM with polynomial kernel for CAD and MI samples classification. }\label{fig:AUCROC}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../../ML/Pics/AUC Conf. matrix}

}

\caption{Confusion matrix on the test set for (A) SVM with RBF kernel for healthy and not-healthy samples classification. (B) Logistic regression model for CAD and MI samples classification. (C) SVM with polynomial kernel for CAD and MI samples classification.}\label{fig:AUCCF}
\end{figure}

\hypertarget{second-layer-for-isolation-of-mi-samples-from-cad-samples}{%
\paragraph{Second layer for isolation of MI samples from CAD
samples}\label{second-layer-for-isolation-of-mi-samples-from-cad-samples}}

To find the best model for training the best set, different models were
trained using their pre-set values. Their AUC-ROC and accuracy results
on the test set are presented in Figure @ref(fig:AUCmodels). The best
model from the AUC-ROC point-of-view was the SVM with linear kernel and
from the accuracy point-of-view, it was the SVM model with an RBF
kernel. For the SVM-linear model the AUC-ROC and accuracy were 0.93 and
0.82, respectively; and for the SVM-RBF, the values were 0.92 and 0.84,
respectively. Both models were hyper-tuned and the ROC curve for their
best performance is presented in Figure @ref(fig:AUCROC)B and C. The
AUC-ROC and accuracy for the SVM-linear model modified to 0.92 and 0.88,
respectively. For the SVM-RBF, these values increased to 0.96 and 0.94,
respectively (Table @ref(tab:AUCML)). The sensitivity for the SVM-linear
and SVM-RBF models were 0.91 and 0.97, respectively; and the specificity
for them was 0.79 and 0.86, respectively. The confusion matrix for both
models is illustrated in Figure @ref(fig:AUCCF)B and C.

\begin{figure}

{\centering \includegraphics[width=0.65\linewidth]{../../ML/Pics/Models AUC 5 mirs CV}

}

\caption{Area under the receiver operating characteristic curve and accuracy of different models trained with AUC-selected miRNAs.}\label{fig:AUCmodels}
\end{figure}

\begin{table}

\caption{\label{tab:AUCML}AUC-ROC and accuracy for SVM with the linear kernel as the best model trained with miRNAs selected based on their individual AUC-ROC on the train and test set before and after hyper-tuning}
\centering
\begin{tabu} to \linewidth {>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
\toprule
\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Pre-set parameters} & \multicolumn{2}{c}{Hyper-tuned} \\
\cmidrule(l{3pt}r{3pt}){3-4} \cmidrule(l{3pt}r{3pt}){5-6}
Model & Metrics & Train & Test & Train & Test\\
\midrule
SVM-linear & AUC-ROC & 0.91 & 0.93 & 0.93 & 0.92\\
 & Accuracy & 0.85 & 0.82 & 0.90 & 0.88\\
SVM-RBF & AUC-ROC & 0.90 & 0.92 & 0.96 & 0.96\\
 & Accuracy & 0.86 & 0.84 & 0.96 & 0.94\\
\bottomrule
\end{tabu}
\end{table}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The prevalence of MI can lead to high-rate mortality in the clinical
setting. However, early diagnosis and application of suitable treatment
protocols can reduce mortality and improve AMI prognosis
({``Cardiovascular Diseases ({CVDs})''} n.d.; Thygesen et al. 2018; Tsao
et al. 2022). Studies have suggested that changes in miRNA expression
may play a significant role in the progression of MI and the subsequent
remodeling (Laggerbauer and Engelhardt 2022). It is believed that the
expression of miRNAs is altered during the various biological processes
correlated with MI within the myocardium or other related tissues (Khan,
Gupta, and Mahapatra 2022). Although several research has been
concentrated on examining free circulating miRNAs in the serum samples
for the detection of cardiac tissue injuries (Kaur et al. 2020), more
information is needed to fully comprehend the miRNAs found in different
blood sub-components like plasma, platelets, and PBMCs. Based on
previous data, PBMCs are critically involved in plaque destabilization
and rupture as well as early inflammatory responses during MI (Mosallaei
et al. 2022; Hapke et al. 2022). Moreover, PBMCs have specific miRNA
profile that is altered under certain pathological conditions which are
great candidates as disease biomarkers (Mosallaei et al. 2022).

PBMCs can respond to several insulting conditions such as MI in the
least possible time with prominent changes in their miRNA profile
(Mosallaei et al. 2022). Considering the regulatory roles, subtle
changes in the transcription of miRNAs can be monitored even before
alteration in the levels of mRNAs and proteins (Schulte et al. 2020).
These features make the miRNAs an early-stage valid diagnostic tool for
the detection of minor and major cell injuries. To date, few studies
have been performed to compare the miRNA profiles in PBMCs belonging to
acute MI patients and other CADs and healthy samples to find a robust
set of identical miRNAs to differentiate these pathological conditions.

In this study, we combined three GEO datasets of healthy, CAD, and MI
samples. Having these samples set alongside bioinformatics analysis and
ML means, it is possible to identify potential biomarker sets and also
effective therapeutic targets. The results of the DEG analysis (Table
@ref(tab:DEGstab) and Figure @ref(fig:venn)) are proof of the close
relationship between the MI and CAD samples. Interestingly, functional
enrichment analysis demonstrated that DEGs in both healthy/CAD and
healthy/MI were strongly correlated to immune cell response which is a
major cellular part of PBMCs. Here, two different sets of miRNAs were
used as biomarker sets for sample classification. miR-21; miR-32; and
miR-186 were selected as differentially expressed miRNAs, and miR-186,
miR-21; miR-29a; miR-197; and miR-296 were selected according to their
AUC-ROC values. As shown in Figure @ref(fig:miRROC), all miRNAs selected
with both approaches had AUC-ROC over 0.9 for the isolation of healthy
and not-healthy samples except for miR-296 and miR-29a. Data confirmed
that the real challenge is to classify CAD and MI samples because of
close overlap. Of 6 miRNAs under investigation in both approaches except
for miR-32, all miRNAs had an AUC-ROC over 0.8 for the discrimination of
CAD and MI samples. Besides, the high AUC-ROC values of miRNAs confirms
their high potential as biomarkers.

ML models when trained with miRNA sets selected by both DEG and AUC-ROC
approaches showed better performance in the classification than each
miRNA. To avoid unwanted complexity and poor predictive values, a
two-layer architecture was also designed. The first layer was for the
discrimination of healthy from not-healthy samples, and the second layer
separated the CAD from MI candidates. As expected, in both approaches a
hyper-tuned SVM model could flawlessly separates healthy from
not-healthy samples using distinct miRNAs sets. The ML models were also
capable of effectively separating CAD from MI patients. Although both
miRNA sets had nearly the same AUC-ROC with their best model, the
accuracy, sensitivity, and specificity were different. The model trained
with AUC-selected miRNAs had better performance in all predictive
values.

Numerous studies have reported different biological processes can affect
the expression of miRNAs in PBMCs. However, there are still
controversies regarding the exact role of miRNAs in the function of
immune cells and the correlation of specific pathological conditions
with miRNA profiles. Several studies have proved the activation of
specific miRNA types in PBMCs under cardiovascular events (S. Li et al.
2015; Yao et al. 2016; Liu et al. 2017; Horita, Farquharson, and Stephen
2021; Bhansali et al. 2022). For instance, there is evidence that the
elevation of miR-186 suppresses the expression of
Cystathionine-\(\gamma\)-lyase, leading to the subsequent secretion of
pro-inflammatory cytokines and cellular lipid accumulation. Besides,
macrophage-derived miR-186 may promote atherosclerotic plaques (Yao et
al. 2016). In line with this claim, we found that miR-186 is
up-regulated in both CAD and MI candidates related to control
counterparts. Surprisingly, the obtained data indicated that the
expression of miR-186 is higher in CAD patients in comparison to MI
(Figure @ref(fig:DEMexp)), To be specific, miR-186 is the only
differentially expressed miRNA between CAD and MI, with a clear
up-regulation in CAD, indicating its main role in the promotion of
atherosclerosis.

As mentioned before, miR-21 was also up-regulated in both MI and CAD
patients in comparison to healthy controls. Moreover, the expression
value of miR-21 was significantly higher in MI than that of the CAD
group (Table @ref(tab:mirExptable)). It is thought that the
up-regulation of miRNA-21 in PBMCs is a compensatory reaction to reduce
T\(_{reg}\) lymphocyte number in response to the reduction of
TGF\(\beta1\) secretion into the plasma through a
TGF\(\beta1\)/smad-independent pathway. In line with previous and
present data, miR-21 can modulate the activity of PBMCs following the
occurrence of cardiovascular diseases (S. Li et al. 2015).

Recent data have supported the elevation of miR-32 in CAD patients with
the calcification of coronary artery. Of note, miR-32 stimulates the
calcification of mouse vascular smooth muscle through the regulation of
bone morphogenetic protein-1, runt-related transcription factor-2
(RUNX2), osteopontin, and bone-specific phosphoprotein matrix GLA
protein (Liu et al. 2017). Likewise, there are some reports associated
with the activity of PBMC miR-32 under several pathologies (Zeng et al.
2021; Wang et al. 2020). The exact role of PBMC miR-32 after
cardiovascular events remained to be elucidated.

Molecular analyses have indicated the regulatory role of miRNAs selected
by the AUC-ROC approach in PBMCs after a cardiovascular event. Two
common miRNAs in DEGs and AUC-ROC approachs, miR-21 and miR-186, were
covered in DEGs miRNA set. Based on numerous reports mir-29a can be
activated in different diseases (Horita, Farquharson, and Stephen
2021).Data analysis indicated that miR-29a is significantly up-regulated
in CAD patients in comparison to healthy and MI groups (Table
@ref(tab:mirExptable)). Increased miR-29a is associated with the
progression of atherosclerosis, and the combination of miR-29a and
ox-LDL was offered as a valid biomarker set for paraclinical
classification (Huang et al. 2016). However, the role of miR-29a in the
function of PBMCs in CAD patients has not been completely examined.

Data indicated that miR-197 is also significantly up-regulated in both
CAD/healthy and MI/healthy groups. In previous studies, it has been
shown that miRNA-197 may play a critical role in regulating the
anti-inflammatory response of IL-35 by affecting pro/ anti-inflammatory
cytokine secretion, M1/M2 macrophage ratio, T\(_{reg}\) lymphocyte
proliferation, and T cell suppression suggesting the potential
diagnostic role of miR-197 in adverse cardiovascular events (Bhansali et
al. 2022).

The findings of this study demonstrate the potential of miR-296 as a
biomarker with high discriminatory power for distinguishing between
samples from individuals with MI and CAD. MiR-296 has been identified as
a key regulator in the development and advancement of atherosclerosis by
controlling the expression of target genes associated with various
biological processes including angiogenesis, inflammation, cholesterol
metabolism, hypertension, cellular proliferation, and apoptosis (H. Li
et al. 2018). In a study it has been shown that miR-296 expression
levels are significantly increased in the PBMCs of CAD patients compared
to healthy controls, suggesting its involvement in regulating
pro-inflammatory cytokines like IL-6 and TNF-a (Fard et al. 2020). These
findings suggest that miR-296 may have a significant impact on the
pathogenesis of atherosclerosis and could potentially serve as a
diagnostic biomarker for CAD or MI.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

In summary, we derived a set of miRNA biomarkers by comparing MI samples
to both healthy and CAD samples. We found that the SVM model performed
best in both the first layer, which separated healthy and not-healthy
samples, and the second layer, which classified MI/CAD samples. The set
of miRNAs selected based on their AUC-ROC values had better performance
in the second layer. Overall, our two-layer structure achieved a
accuracy of 0.96. This demonstrates the potential for combining
bioinformatics and machine learning techniques to identify novel
biomarkers and gain a deeper understanding of myocardial infarction.

\hypertarget{references}{%
\section{References}\label{references}}

\begin{align}
a^2+b^2=c^2
\end{align}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-197}{}}%
Bhansali, Shipra, Amit Kumar Yadav, Chetan Bakshi, and Veena Dhawan.
2022. {``Interleukin-35 {Mitigates} Ox-{LDL}-{Induced} {Proatherogenic}
{Effects} via {Modulating} {miRNAs} {Associated} with {Coronary}
{Artery} {Disease} ({CAD}).''} \emph{Cardiovascular Drugs and Therapy},
April. \url{https://doi.org/10.1007/s10557-022-07335-x}.

\leavevmode\vadjust pre{\hypertarget{ref-75}{}}%
Canali, Raffaella, Lucia Natarelli, Guido Leoni, Elena Azzini, Raffaella
Comitato, Oezgur Sancak, Luca Barella, and Fabio Virgili. 2014.
{``Vitamin {C} Supplementation Modulates Gene Expression in Peripheral
Blood Mononuclear Cells Specifically Upon an Inflammatory Stimulus: A
Pilot Study in Healthy Subjects.''} \emph{Genes \& Nutrition} 9 (3):
390. \url{https://doi.org/10.1007/s12263-014-0390-x}.

\leavevmode\vadjust pre{\hypertarget{ref-CVD}{}}%
{``Cardiovascular Diseases ({CVDs}).''} n.d. Accessed March 12, 2023.
\url{https://www.who.int/news-room/fact-sheets/detail/cardiovascular-diseases-(cvds)}.

\leavevmode\vadjust pre{\hypertarget{ref-296a}{}}%
Fard, Toktam Kazemi, Samareh Tavakoli, Reza Ahmadi, Nariman Moradi, Reza
Fadaei, Asghar Mohammadi, and Soudabeh Fallah. 2020. {``Evaluation of
{Ip10} and {miRNA} 296-a {Expression} {Levels} in {Peripheral} {Blood}
{Mononuclear} {Cell} of {Coronary} {Artery} {Disease} {Patients} and
{Controls}.''} \emph{DNA and Cell Biology} 39 (9): 1678--84.
\url{https://doi.org/10.1089/dna.2020.5650}.

\leavevmode\vadjust pre{\hypertarget{ref-PBMC-miR}{}}%
Gao, Jie, Jia Liu, Ying Zhang, BaoYi Guan, Hua Qu, Hua Chai, WenTing
Wang, XiaoJuan Ma, and DaZhuo Shi. 2020. {``{PBMCs}-{Derived} {microRNA}
{Signature} as a {Prethrombotic} {Status} {Discriminator} in {Stable}
{Coronary} {Artery} {Disease}.''} \emph{Thrombosis and Haemostasis} 120
(01): 121--31. \url{https://doi.org/10.1055/s-0039-1700518}.

\leavevmode\vadjust pre{\hypertarget{ref-Meder23}{}}%
Hapke, Nils, Margarete Heinrichs, DiyaaElDin Ashour, Elena Vogel, Ulrich
Hofmann, Stefan Frantz, and Gustavo Campos Ramos. 2022.
{``Identification of a Novel Cardiac Epitope Triggering {T}-Cell
Responses in Patients with Myocardial Infarction.''} \emph{Journal of
Molecular and Cellular Cardiology} 173 (December): 25--29.
\url{https://doi.org/10.1016/j.yjmcc.2022.09.001}.

\leavevmode\vadjust pre{\hypertarget{ref-numpy}{}}%
Harris, Charles R., K. Jarrod Millman, Stéfan J. van der Walt, Ralf
Gommers, Pauli Virtanen, David Cournapeau, Eric Wieser, et al. 2020.
{``Array Programming with {NumPy}.''} \emph{Nature} 585 (7825): 357--62.
\url{https://doi.org/10.1038/s41586-020-2649-2}.

\leavevmode\vadjust pre{\hypertarget{ref-scikitopt}{}}%
Head, Tim, Manoj Kumar, Holger Nahrstaedt, Gilles Louppe, and Iaroslav
Shcherbatyi. 2021. \emph{Scikit-Optimize/Scikit-Optimize} (version
v0.9.0). Zenodo. \url{https://doi.org/10.5281/zenodo.5565057}.

\leavevmode\vadjust pre{\hypertarget{ref-29aa}{}}%
Horita, Masahiro, Colin Farquharson, and Louise A Stephen. 2021. {``The
Role of {miR}‐29 Family in Disease.''} \emph{Journal of Cellular
Biochemistry} 122 (7): 696--715.
\url{https://doi.org/10.1002/jcb.29896}.

\leavevmode\vadjust pre{\hypertarget{ref-29a}{}}%
Huang, Yu-Qing, An-Ping Cai, Ji-Yan Chen, Cheng Huang, Jie Li, and
Ying-Qing Feng. 2016. {``The {Relationship} of {Plasma} {miR}-29a and
{Oxidized} {Low} {Density} {Lipoprotein} with {Atherosclerosis}.''}
\emph{Cellular Physiology and Biochemistry} 40 (6): 1521--28.
\url{https://doi.org/10.1159/000453202}.

\leavevmode\vadjust pre{\hypertarget{ref-m2}{}}%
Kalayinia, Samira, Fateme Arjmand, Majid Maleki, Mahshid Malakootian,
and Chandra Pal Singh. 2021. {``{MicroRNAs}: Roles in Cardiovascular
Development and Disease.''} \emph{Cardiovascular Pathology} 50
(January): 107296. \url{https://doi.org/10.1016/j.carpath.2020.107296}.

\leavevmode\vadjust pre{\hypertarget{ref-11-15}{}}%
Kaur, Amanpreet, Sharon T Mackin, Kenny Schlosser, Fui Lin Wong, Malik
Elharram, Christian Delles, Duncan J Stewart, Natalie Dayan, Tara
Landry, and Louise Pilote. 2020. {``Systematic Review of {microRNA}
Biomarkers in Acute Coronary Syndrome and Stable Coronary Artery
Disease.''} \emph{Cardiovascular Research} 116 (6): 1113--24.
\url{https://doi.org/10.1093/cvr/cvz302}.

\leavevmode\vadjust pre{\hypertarget{ref-24}{}}%
Khan, Abrar A., Vinayak Gupta, and Nitish R. Mahapatra. 2022. {``Key
Regulatory {miRNAs} in Lipid Homeostasis: {Implications} for
Cardiometabolic Diseases and Development of Novel Therapeutics.''}
\emph{Drug Discovery Today} 27 (8): 2170--80.
\url{https://doi.org/10.1016/j.drudis.2022.05.003}.

\leavevmode\vadjust pre{\hypertarget{ref-23}{}}%
Laggerbauer, Bernhard, and Stefan Engelhardt. 2022. {``{MicroRNAs} as
Therapeutic Targets in Cardiovascular Disease.''} \emph{Journal of
Clinical Investigation} 132 (11): e159179.
\url{https://doi.org/10.1172/JCI159179}.

\leavevmode\vadjust pre{\hypertarget{ref-BER}{}}%
Lazar, C., S. Meganck, J. Taminau, D. Steenhoff, A. Coletta, C. Molter,
D. Y. Weiss-Solis, R. Duque, H. Bersini, and A. Nowe. 2013. {``Batch
Effect Removal Methods for Microarray Gene Expression Data Integration:
A Survey.''} \emph{Briefings in Bioinformatics} 14 (4): 469--90.
\url{https://doi.org/10.1093/bib/bbs037}.

\leavevmode\vadjust pre{\hypertarget{ref-296}{}}%
Li, Heng, Heng Li, Xin-Ping Ouyang, Ting Jiang, Xi-Long Zheng, Xi-Long
Zheng, Ping-Ping He, and Guo-Jun Zhao. 2018. {``{MicroRNA}-296: A
Promising Target in the Pathogenesis of Atherosclerosis?''}
\emph{Molecular Medicine} 24 (1): 12--12.
\url{https://doi.org/10.1186/s10020-018-0012-y}.

\leavevmode\vadjust pre{\hypertarget{ref-21}{}}%
Li, Sihui, Qian Fan, Shaolin He, Tingting Tang, Yuhua Liao, and
Jiangjiao Xie. 2015. {``{MicroRNA}-21 {Negatively} {Regulates} {Treg}
{Cells} {Through} a {TGF}-\(\beta1\)/{Smad}-{Independent} {Pathway} in
{Patients} with {Coronary} {Heart} {Disease}.''} \emph{Cellular
Physiology and Biochemistry} 37 (3): 866--78.
\url{https://doi.org/10.1159/000430214}.

\leavevmode\vadjust pre{\hypertarget{ref-32-2}{}}%
Liu, Jianghua, Xinhua Xiao, Yingying Shen, Ling Chen, Canxin Xu, Heng
Zhao, Ying Wu, et al. 2017. {``{MicroRNA}-32 Promotes Calcification in
Vascular Smooth Muscle Cells: {Implications} as a Novel Marker for
Coronary Artery Calcification.''} Edited by Yin Tintut. \emph{PLOS ONE}
12 (3): e0174138. \url{https://doi.org/10.1371/journal.pone.0174138}.

\leavevmode\vadjust pre{\hypertarget{ref-67}{}}%
Maciejak, Agata, Marek Kiliszek, Marcin Michalak, Dorota Tulacz,
Grzegorz Opolski, Krzysztof Matlak, Slawomir Dobrzycki, Agnieszka
Segiet, Monika Gora, and Beata Burzynska. 2015. {``Gene Expression
Profiling Reveals Potential Prognostic Biomarkers Associated with the
Progression of Heart Failure.''} \emph{Genome Medicine} 7 (1): 26.
\url{https://doi.org/10.1186/s13073-015-0149-z}.

\leavevmode\vadjust pre{\hypertarget{ref-09}{}}%
Matone, Alice, Colm M. O'Grada, Eugene T. Dillon, Ciara Morris, Miriam
F. Ryan, Marianne Walsh, Eileen R. Gibney, et al. 2015. {``Body Mass
Index Mediates Inflammatory Response to Acute Dietary Challenges.''}
\emph{Molecular Nutrition \& Food Research} 59 (11): 2279--92.
\url{https://doi.org/10.1002/mnfr.201500184}.

\leavevmode\vadjust pre{\hypertarget{ref-fRMA}{}}%
McCall, M. N., B. M. Bolstad, and R. A. Irizarry. 2010. {``Frozen Robust
Multiarray Analysis ({fRMA}).''} \emph{Biostatistics} 11 (2): 242--53.
\url{https://doi.org/10.1093/biostatistics/kxp059}.

\leavevmode\vadjust pre{\hypertarget{ref-Barcode}{}}%
McCall, Matthew N., Karan Uppal, Harris A. Jaffee, Michael J. Zilliox,
and Rafael A. Irizarry. 2011. {``The {Gene} {Expression} {Barcode}:
Leveraging Public Data Repositories to Begin Cataloging the Human and
Murine Transcriptomes.''} \emph{Nucleic Acids Research} 39 (suppl\_1):
D1011--15. \url{https://doi.org/10.1093/nar/gkq1259}.

\leavevmode\vadjust pre{\hypertarget{ref-pandas}{}}%
McKinney, Wes. 2010. {``{D}ata {S}tructures for {S}tatistical
{C}omputing in {P}ython.''} In \emph{{P}roceedings of the 9th {P}ython
in {S}cience {C}onference}, edited by Stéfan van der Walt and Jarrod
Millman, 56--61.
\href{https://doi.org/\%2010.25080/Majora-92bf1922-00a\%20}{https://doi.org/
10.25080/Majora-92bf1922-00a }.

\leavevmode\vadjust pre{\hypertarget{ref-Meder6}{}}%
Mosallaei, Meysam, Naeim Ehtesham, Shima Rahimirad, Mostafa Saghi, Nasim
Vatandoost, and Sharifeh Khosravi. 2022. {``{PBMCs}: A New Source of
Diagnostic and Prognostic Biomarkers.''} \emph{Archives of Physiology
and Biochemistry} 128 (4): 1081--87.
\url{https://doi.org/10.1080/13813455.2020.1752257}.

\leavevmode\vadjust pre{\hypertarget{ref-scikit-learn}{}}%
Pedregosa, F., G. Varoquaux, A. Gramfort, V. Michel, B. Thirion, O.
Grisel, M. Blondel, et al. 2011. {``Scikit-Learn: Machine Learning in
{P}ython.''} \emph{Journal of Machine Learning Research} 12: 2825--30.

\leavevmode\vadjust pre{\hypertarget{ref-ML}{}}%
Reel, Parminder S., Smarti Reel, Ewan Pearson, Emanuele Trucco, and
Emily Jefferson. 2021. {``Using Machine Learning Approaches for
Multi-Omics Data Analysis: {A} Review.''} \emph{Biotechnology Advances}
49 (July): 107739.
\url{https://doi.org/10.1016/j.biotechadv.2021.107739}.

\leavevmode\vadjust pre{\hypertarget{ref-miR}{}}%
Schulte, Christian, Temo Barwari, Abhishek Joshi, Tanja Zeller, and
Manuel Mayr. 2020. {``Noncoding {RNAs} Versus {Protein} {Biomarkers} in
{Cardiovascular} {Disease}.''} \emph{Trends in Molecular Medicine} 26
(6): 583--96. \url{https://doi.org/10.1016/j.molmed.2020.02.001}.

\leavevmode\vadjust pre{\hypertarget{ref-m1}{}}%
Schulte, Christian, Mahir Karakas, and Tanja Zeller. 2017.
{``{microRNAs} in Cardiovascular Disease -- Clinical Application.''}
\emph{Clinical Chemistry and Laboratory Medicine (CCLM)} 55 (5).
\url{https://doi.org/10.1515/cclm-2016-0576}.

\leavevmode\vadjust pre{\hypertarget{ref-36}{}}%
Soler-Botija, Carolina, Carolina Gálvez-Montón, and Antoni Bayés-Genís.
2019. {``Epigenetic {Biomarkers} in {Cardiovascular} {Diseases}.''}
\emph{Frontiers in Genetics} 10 (October): 950.
\url{https://doi.org/10.3389/fgene.2019.00950}.

\leavevmode\vadjust pre{\hypertarget{ref-17}{}}%
Thygesen, Kristian, Joseph S. Alpert, Allan S. Jaffe, Bernard R.
Chaitman, Jeroen J. Bax, David A. Morrow, Harvey D. White, and The
Executive Group on behalf of the Joint European Society of Cardiology
(ESC)/American College of Cardiology (ACC)/American Heart Association
(AHA)/World Heart Federation (WHF) Task Force for the Universal
Definition of Myocardial Infarction. 2018. {``Fourth {Universal}
{Definition} of {Myocardial} {Infarction} (2018).''} \emph{Circulation}
138 (20). \url{https://doi.org/10.1161/CIR.0000000000000617}.

\leavevmode\vadjust pre{\hypertarget{ref-ML-Bio}{}}%
Torun, Furkan M., Sebastian Virreira Winter, Sophia Doll, Felix M.
Riese, Artem Vorobyev, Johannes B. Mueller-Reif, Philipp E. Geyer, and
Maximilian T. Strauss. 2023. {``Transparent {Exploration} of {Machine}
{Learning} for {Biomarker} {Discovery} from {Proteomics} and {Omics}
{Data}.''} \emph{Journal of Proteome Research} 22 (2): 359--67.
\url{https://doi.org/10.1021/acs.jproteome.2c00473}.

\leavevmode\vadjust pre{\hypertarget{ref-18}{}}%
Tsao, Connie W., Aaron W. Aday, Zaid I. Almarzooq, Alvaro Alonso, Andrea
Z. Beaton, Marcio S. Bittencourt, Amelia K. Boehme, et al. 2022.
{``Heart {Disease} and {Stroke} {Statistics}---2022 {Update}: {A}
{Report} {From} the {American} {Heart} {Association}.''}
\emph{Circulation} 145 (8).
\url{https://doi.org/10.1161/CIR.0000000000001052}.

\leavevmode\vadjust pre{\hypertarget{ref-32}{}}%
Wang, Dan, Ting Zeng, Zhi Lin, Lu Yan, Fenglin Wang, Lanlan Tang, Leyuan
Wang, Daolin Tang, Pan Chen, and Minghua Yang. 2020. {``Long Non-Coding
{RNA} {Snhg5} Regulates Chemotherapy Resistance Through the
{miR}-32/{Dnajb9} Axis in Acute Myeloid Leukemia.''} \emph{Biomedicine
\& Pharmacotherapy} 123 (March): 109802.
\url{https://doi.org/10.1016/j.biopha.2019.109802}.

\leavevmode\vadjust pre{\hypertarget{ref-186}{}}%
Yao, Yan, Xin Zhang, Hai-peng Chen, Liang Li, Wei Xie, Gang Lan,
Zhen-wang Zhao, Xi-Long Zheng, Zong-bao Wang, and Chao-ke Tang. 2016.
{``{MicroRNA}-186 Promotes Macrophage Lipid Accumulation and Secretion
of Pro-Inflammatory Cytokines by Targeting Cystathionine
\(\gamma\)-Lyase in {THP}-1 Macrophages.''} \emph{Atherosclerosis} 250
(July): 122--32.
\url{https://doi.org/10.1016/j.atherosclerosis.2016.04.030}.

\leavevmode\vadjust pre{\hypertarget{ref-Macro}{}}%
Yap, Jonathan, Jason Irei, Javier Lozano-Gerona, Selena Vanapruks,
Tianmai Bishop, and William A. Boisvert. 2023. {``Macrophages in Cardiac
Remodelling After Myocardial Infarction.''} \emph{Nature Reviews
Cardiology}, January. \url{https://doi.org/10.1038/s41569-022-00823-5}.

\leavevmode\vadjust pre{\hypertarget{ref-magic}{}}%
Zeng, Zl, Qingyun Zhu, Zhibo Zhao, Xuyu Zu, and Jianghua Liu. 2021.
{``Magic and Mystery of {microRNA}‐32.''} \emph{Journal of Cellular and
Molecular Medicine} 25 (18): 8588--8601.
\url{https://doi.org/10.1111/jcmm.16861}.

\end{CSLReferences}

\bibliographystyle{spmpsci}
\bibliography{ForArticleRefrence.bib,discussion.bib,Introduction.bib,mir296.bib}

\end{document}